Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex
more

\begin{definition}\rm
A formula $\varphi$ is in \emph{definitional form with respect to}
$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
form
%
%\kern-\abovedisplayskip
%\kern+\abovedisplayshortskip
%
\[(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi\]
%
where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
%ct added
$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
for $i = 1, \ldots, n$,
%
\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
\jb{do we?}
and $\psi$ contains no function definitions.
We call $\psi$ the \emph{conjecture} of $\varphi$.
%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
% Now fixed above -ajr
\end{definition}

@@

For example, if $T$ is the theory of integer arithmetic,
the set $\{ \forallf{\con{f}} \vec x:\typeint.\;\, \con{f}( x ) \teq 0 \}$ is admissible.
An admissible set can include definitions of nonterminating functions,
\ct{There is no notion of termination in FOL. We either define this properly or we just replace "definitions of nonterminating functions, e.g.," with "definitions like"}
\ajr{I agree. Jasmin, can you fix this?  We need to be careful in the same way when we talk about the "terminating" functions fragment we handle.}
e.g.,
$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\allowbreak\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
On the other hand, the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not admissible,
since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
and there is no model of $T$ satisfying $\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1$.
Consider the set 
$\{ 
 \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;
 \forallf{\con{g}} x.\;\, \lite\bigl( \con{f}( x ) \teq 5,\;
   \con{g}( x ) \teq \con{f}( x ),\,
   \con{g}( x ) \teq \con{g}( x ) + 1)
\}$.
While this set has a model where $\con f$ and $\con g$ are interpreted as the constant $5$, it is not admissible
since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\, \con{g}( x ) \teq \con{f}( x ),\, \con{g}( x ) \teq \con{g}( x ) + 1 )$.

>>>>

\begin{definition}\rm
A formula $\varphi$ is in \emph{definitional form with respect to}
$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
form
%
%\kern-\abovedisplayskip
%\kern+\abovedisplayshortskip
%
\[(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi\]
%
where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
%ct added
$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
for $i = 1, \ldots, n$,
%
\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
\jb{do we? I'd think that the admissibility criterion below is enough.}
and $\psi$ contains no function definitions.
We call $\psi$ the \emph{conjecture} of $\varphi$.
%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
% Now fixed above -ajr
\end{definition}

@@

For example, if $T$ is the theory of integer arithmetic,
the set $\{ \forallf{\con{f}} \vec x:\typeint.\;\, \con{f}( x ) \teq 0 \}$ is admissible.
An admissible set can include definitions of nonterminating functions,
\ct{There is no notion of termination in FOL. We either define this properly or we just replace "definitions of nonterminating functions, e.g.," with "definitions like"}
\ajr{I agree. Jasmin, can you fix this?  We need to be careful in the same way when we talk about the "terminating" functions fragment we handle.}
\jb{Yes, I need to rework and expand this whole part. Today.}
e.g.,
$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\allowbreak\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
On the other hand, the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not admissible,
since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
and there is no model of $T$ satisfying $\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1$.
Consider the set 
$\{ 
 \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;
 \forallf{\con{g}} x.\;\, \lite\bigl( \con{f}( x ) \teq 5,\;
   \con{g}( x ) \teq \con{f}( x ),\,
   \con{g}( x ) \teq \con{g}( x ) + 1)
\}$.
While this set has a model where $\con f$ and $\con g$ are interpreted as the constant $5$, it is not admissible
since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\, \con{g}( x ) \teq \con{f}( x ),\, \con{g}( x ) \teq \con{g}( x ) + 1 )$.
